The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA)

### 2014-05-07 hera-concept2013.tex
update...

%% copernicus_discussions.cls
%\documentclass[ars, hvmath, online]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus_discussions}

\def\kperp{k_{\bot}}
\def\kpar{k_{\|}}
\def\nwnh{{\sl NWNH}}

@@

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged project
that uses the unique properties of the 21-cm line of neutral hydrogen to probe the
Epoch of Reionization (EoR) and the preceding Dark Ages. During these epochs, roughly
0.3-1 Gyr after the Big Bang, the first stars and black holes heated and reionized
the Universe, which at that time existed as a nearly uniform sea of warm neutral
hydrogen.   This period of cosmic reionization represents one of the few global
phase changes in the evolution of the Universe.
Direct observation of the large scale structure of reionization and its
evolution with time will have a profound impact on our understanding of the birth of
the first galaxies and black holes, their influence on the intergalactic medium
(IGM), and cosmology.

Figure \ref{fig:theUniverse} is a cartoon showing cosmic evolution from the Big Bang on the left through
to HERA today.

@@

HERA overall is a phased set of experiments that build on its predecessors and preceding stages.
The first phase of the HERA roadmap entailed the operation of the
PAPER and MWA telescopes to explore techniques and designs required to
detect the primordial HI signal in the presence of radio continuum
foreground emission some four orders of magnitude brighter. Studies
with PAPER and the MWA have led to a new understanding of the
interplay of foreground and instrumental systematics in the context of
a three-dimensional cosmological intensity-mapping experiment.  YYYYYGIVE REFSYYYYY 
We are
now able to remove foregrounds to the limits of our sensitivity with
these instruments, culminating in the first physically meaningful
upper limits on the power spectrum of 21~cm emission from
reionization (see Fig \ref{fig:eor_pspec}).

Building on this understanding, the next stage of HERA entails a new
14m diameter antenna element that is optimized both for sensitivity
and for minimizing foreground systematics.  Arranging these elements
in a compact hexagonal grid yields an array that facilitates
calibration, leverages proven foreground removal techniques, and is
scalable to large collecting areas. The array will be located near the current PAPER
experiment
in the radio quiet environment of the SKA site in Karoo, South Africa,
and will have a sensitivity close to two orders of magnitude better than
PAPER and the MWA.  HERA's sensitivity coupled with broader frequency coverage, 
will enable HERA to paint an
uninterrupted picture through reionization, back to the end of the
Dark Ages.  As an array, it can be used in smaller arrays along the way to 
achieve timely science.  Two benchmarks are planned:

@@

This paper will present a summary of the current understanding
of the signal characteristics and measurements and describe this planned
HERA telescope to be built to detect and characterize the EoR power
spectrum.

\section{Background}
\label{sec:background}
\subsection{Hubble's Law}
\label{sec:hubble}
Modern cosmology has provided a spectacular understanding of the cosmic history of
our Universe starting from the instantaneous moment after the explosion of our
spacetime that we call ``the Big Bang'', which occurred about 13.8 Gyr ago, to the
advent of the large structures that seeded our cosmic current structure. After the
Big Bang, matter and space went streaming outward and is manifest to us by the simple
expression called Hubble's Law:
\begin{equation}
v = H_od
\end{equation}
where $v$ is the velocity of the object away from us, $d$ is the distance to the
object $H_o$ is the so-called ``Hubble Constant'', which has a value of 68 Mpc/km/sec
\ref{planck}, in astronomer's units, where Mpc denotes megaparsecs, where a parsec is
$3.086 \times 10^13$ km. Using measured Doppler shifts one may measure $v$, so if one
knows $H_o$ one may directly compute the distance to the object. Velocity is
typically given as a ``redshift'', denoted as $z$, related to one another by the
non-relativistic Doppler shift equation $z \sim v/c$, where $c$ is the speed of light
in vacuo. And then by Hubble's Law, that redshift conveys a distance.

\subsection{Power Spectrum}
\label{sec:pspec}
One tool that is typically employed in studying large scale parameters of the
Universe, is to collapse the richness of its spatial variation to its spatial Fourier
transform and compute a power spectrum of this. It is the parameters of this power
spectrum that may be predicted by cosmology, rather than the essentially random
spatial variation about it. It is analogous to the fact that one may get a few broad
parameters of an FM signal by Fourier transforming its time series, regardless of whether it is
Beethoven or M\"{o}tley Cr\"{u}e, which have very different values as a function of time.

Cosmology typically encodes this data as power in spherical harmonics.  Epoch of
Reionization studies typically use wavevectors, $\k$.  This 3-D vector may be decomposed
into plane-of-sky values (${\bf \kperp}$) and a line-of-sight term (${\kpar}$).  The overall
wavevector may then be written as $\k = {\bf \kperp} + \kpar{\bf z}$.  As a wavevector,
the units of $\k$ are 1/Mpc (inverse length).  Note that although modern cosmology believes
it has pegged the value of Hubble's constant, to hedge their bets, they will often factor out
$h=H_o/100$ as an overall scale adjustment, such that the units are $h$/Mpc.  The units
of the power spectrum magnitudes are given in volume-adjusted power-squared (mK$^2$), denoted
$\Delta ^2$.

It is also interesting to note that another ``foreground'' from our perspective is the Cosmic Microwave Background.

@@

@@

\subsection{Frequency Range}
The rest frequency ($\nu_o$) of the hyperfine transition of neutral hydrogen is at
1420.4 MHz. This frequency gets red-shifted to lower frequencies in deep astronomical
observations due to the expansion of the Universe, as discussed above. This is
characterized by the red-shift parameter $z$ as
\begin{equation}
\centering
\label{eq:redshift}
\nu_s(z) = \nu_o/(1+z)
\end{equation}
The age of the Universe at a particular redshift is given by the cosmological model
and a key goal of cosmology is to determine those parameters, however existing models
allow us to determine these values to our desired accuracy. Figure
\ref{fig:theUniverse} plots the red-shifted frequency as a function of red-shift (the
blue curve), with the upper axis showing the cosmic evolution time since the Big Bang
(using parameters from REF PLANCK). The bandwidth for HERA is demarcated by the
horizontal dashed lines. The black dashed lines denote the ``EOR'' core bandwidth,
and the extension to the blue dotted lines is to capture earlier epochs (going lower
in frequency) and find the post-EOR ``null'' (going higher in frequency) as discussed
below. Table \ref{tab:heraband} provides the details of how the frequencies map to
redshift and cosmic time.

@@

\subsection{Delay}
To exploit the smooth spectral response filtering of the foregrounds

\subsection{Sensitivity Metric}
The equation for sensitivity has been described in previous works [ref ] and is reproduced below:
\begin{equation}
\centering
\label{eq:sensitivity}
\Delta^2_N(k) 
\end{equation}
Note that $\Delta^2_N(k)$ is the standard radiometric sensitivity equation, scaled by
the volume in $k$-space, normalized by the power spectrum Fourier coefficient, and
reduced by the number of independent samples in a given $k$-mode bin, which may have
both coherent and incoherent application.

Why close-packed.  Hexagonal has 60degree symmetry, not 90degree symmetry like for rectangular.

Why 14-m dishes.

\section{System Overview}
\subsection{Antenna}
14-m

\subsection{Analog System}
amplifiers and stuff

\subsection{Digital System}
fpga and gpus and stuff

\section{Site}
Karoo

>>>>

%% copernicus_discussions.cls
%\documentclass[ars, hvmath, online]{/Users/daviddeboer1/Documents/Papers/Copernicus_LaTeX_Package_v_2_7/copernicus_discussions}
\usepackage{amsmath}
\def\kperp{k_{\bot}}
\def\kpar{k_{\|}}
\def\nwnh{{\sl NWNH}}

@@

\introduction  %% \introduction[modified heading if necessary]
\label{sec:intro}
The Donald C. Backer Hydrogen Epoch of Reionization Array (HERA) is a staged project
that uses the unique properties of the 21-cm line of neutral hydrogen to probe the
Epoch of Reionization (EoR) and the preceding Dark Ages. During these epochs, roughly
0.3-1 Gyr after the Big Bang, the first stars and black holes heated and reionized
the Universe, which at that time existed as a nearly uniform sea of warm neutral
hydrogen.  Figure \ref{fig:theUniverse} is a cartoon showing cosmic evolution from the Big Bang on the left through
to HERA today. The period of cosmic reionization represents one of the few global
phase changes in the evolution of the Universe.
Direct observation of the large scale structure of reionization and its
evolution with time will have a profound impact on our understanding of the birth of
the first galaxies and black holes, their influence on the intergalactic medium
(IGM), and cosmology.

@@

The HERA roadmap is a phased set of experiments that build on its predecessors and preceding stages.
The first phase of the HERA roadmap entailed the operation of the
PAPER and MWA telescopes to explore techniques and designs required to
detect the primordial HI signal in the presence of radio continuum
foreground emission some four orders of magnitude brighter. Studies
with PAPER and the MWA have led to a new understanding of the
interplay of foreground and instrumental systematics in the context of
a three-dimensional cosmological intensity-mapping experiment.  YYYYYGIVE REFSYYYYY 
We are
now able to remove foregrounds to the limits of our sensitivity with
these instruments, culminating in the first physically meaningful
upper limits on the power spectrum of 21~cm emission from
reionization (see Fig \ref{fig:eor_pspec}).

Building on this understanding, the next stage of HERA incorporates a new
14m diameter antenna element that is optimized both for sensitivity
and for minimizing foreground systematics.  Arranging these elements
in a compact hexagonal grid yields an array that facilitates
calibration, leverages proven foreground removal techniques, and is
scalable to large collecting areas. The array will be located near the current PAPER
experiment
in the radio quiet environment of the SKA site in Karoo, South Africa,
and will have a sensitivity close to two orders of magnitude better than
PAPER and the MWA.  HERA's sensitivity coupled with broader frequency coverage, 
will enable HERA to paint an
uninterrupted picture through reionization, back to the end of the
Dark Ages.  As an array, it can be used in smaller arrays along the way to 
achieve timely science.  Two benchmarks are planned:

@@

This paper will present a summary of the current understanding
of the signal characteristics and measurements and describe this planned
HERA telescope to be built to detect and characterize the EoR power
spectrum.  After presenting the scientific and motivation and background, the
derived specificiations and resulting system will be described.  The data pipeline
will then be summarized, follwed by the conclusion.

@@

\section{Background}
\label{sec:background}
\subsection{Hubble's Law}
\label{sec:hubble}
Modern cosmology has provided a spectacular understanding of the cosmic history of
our Universe starting from the instantaneous moment after the explosion of our
spacetime that we call ``the Big Bang'', which occurred about 13.8 Gyr ago, to the
advent of the large structures that seeded our cosmic current structure. After the
Big Bang, matter and space went streaming outward and is manifest to us by the simple
expression called Hubble's Law:
\begin{equation}
v = H_od
\end{equation}
where $v$ is the velocity of the object away from us, $d$ is the distance to the
object $H_o$ is the so-called ``Hubble Constant'', which has a value of 68 Mpc/km/sec
\ref{planck}, in astronomer's units, where Mpc denotes megaparsecs, where a parsec is
$3.086 \times 10^13$ km. Using measured Doppler shifts one may determine $v$, so if one
knows $H_o$ one may directly compute the distance to the object. Velocity is
typically given as a ``redshift'', denoted as $z$, related to one another by the
non-relativistic Doppler shift equation $z \sim v/c$, where $c$ is the speed of light
in vacuo. And then by Hubble's Law, that redshift conveys a distance.

\subsection{Power Spectrum}
\label{sec:pspec}
One tool that is typically employed in studying large scale parameters of the
Universe, is to collapse the richness of its spatial variation to its spatial Fourier
transform and compute a power spectrum of this. It is the parameters of this power
spectrum that may be predicted by cosmology, rather than the essentially random
particular spatial variation about it. It is analogous to the fact that one may get a few broad
parameters of an FM signal by Fourier transforming its time series, regardless of whether it is
Beethoven or M\"{o}tley Cr\"{u}e, which have very different values as a function of time.

Cosmology typically encodes this data as power in spherical harmonics.  Epoch of
Reionization studies typically use wavevectors, $\k$.  This 3-D vector may be decomposed
into plane-of-sky values (${\bf \kperp}$) and a line-of-sight term (${\kpar}$).  The overall
wavevector may then be written as $\k = {\bf \kperp} + \kpar{\bf z}$.  As a wavevector,
the units of $\k$ are 1/Mpc (inverse length).  Note that although modern cosmology believes
it has pegged the value of Hubble's constant, to hedge their bets, they will often factor out
$h=H_o/100$ as an overall scale adjustment, such that the wavevector ``units'' are $h$/Mpc.  The units
of the power spectrum magnitudes are given in $\k$-volume-adjusted power-squared (mK$^2$), denoted
$\Delta ^2$.

Note that the initial approach to detecting and characterizing the EOR is therefore not imaging, but 
to detect power in a few of the power spectrum bins.  One 
may exploit this to get more samples of the power in specific bins to beat down the noise, at the expense of 
imaging, although for very compact arrays there is less of an imaging penalty as more of the bins
are filled within a radius of support.  The implications of the configuration to the array sensitivity
will be touched upon below.

It is also interesting to note that another ``foreground'' from our perspective is the Cosmic Microwave Background.  
In this case, the detectable signal is the difference in flux between the hydrogren and the CMB.  It acts a bit like the
weatherman's "green screen", where anything in a specfic shade of green is invisible.

\subsection{Filtering}
The wedge etc

@@

\subsection{Frequency Range}
The rest frequency ($\nu_o$) of the hyperfine transition of neutral hydrogen is at
1420.4 MHz. This frequency gets red-shifted to lower frequencies in deep astronomical
observations due to the expansion of the Universe, as discussed above. This is
characterized by the red-shift parameter $z$ as
\begin{equation}
\centering
\label{eq:redshift}
\nu_s(z) = \nu_o/(1+z)
\end{equation}
The age of the Universe at a particular redshift is given by the cosmological model
and a key goal of cosmology is to determine those parameters, however existing models
allow us to determine these values to our desired accuracy. Figure
\ref{fig:theUniverse} plots the red-shifted frequency as a function of red-shift (the
blue curve), with the upper axis showing the cosmic evolution time since the Big Bang
(using parameters from REF PLANCK). The bandwidth for HERA is demarcated by the
horizontal dashed lines. The black dashed lines denote the ``EOR'' core bandwidth,
and the extension to the blue dotted lines is to capture earlier epochs (going lower
in frequency) and find the post-EOR ``null'' (going higher in frequency) as discussed
below. Figure \ref{fig:theUniverse} shows the mapping.

@@

\subsection{Delay and Systematics}
To exploit the smooth spectral response filtering of the foregrounds

\subsection{Sensitivity}
The sensitivity specification is to maximize performance per cost, which determines the element 
diameter ($D$) and the number of elements ($N$) subject to the constraints above.  One therefore 
needs a model of cost and performance as a function of $D$ and $N$.  Given a fairly mature element
and system design, a bottoms-up costing appropriate for element diameters from about 6-m to 20-m 
has been done for hex-numbers corresponding to element counts from 37 to 631.  The costs here
are only those associated with delivering the array to that scope on-site, so excluding development and
science.

The equation for sensitivity has been described in previous works [REFS] and depends on many parameters
related to the instrumentation, the configuration, the location, the observing strategy, etc.  A useful form for 
the proposed compact configuration is Eq 25 in Parsons et al 2012 and reproduced here:
\begin{equation}
\centering
\label{eq:sensitivity}
\begin{split}
\Delta^2_N(k) \approx 60 \left[\frac{k}{0.1h\text{Mpc}^{-1}}\right]^\frac{5}{2}
                                         \left[\frac{6 \text{MHz}}{B}\right]^\frac{1}{2}
                                         \left[\frac{1}{\Delta\ln k}\right]^\frac{1}{2} \\
                        \times       \left[\frac{\Omega}{0.76\text{sr}}\right]
                                         \left[\frac{T_\text{sys}}{500 \text{K}}\right]^2
                                         \left[\frac{6 \text{hrs}}{t_\text{per\_day}}\right]^\frac{1}{2} \\
                        \times       \left[\frac{120\text{days}}{t}\right]
                                         \left[\frac{32}{N}\right]
                                         \left[\frac{\text{10}^4f_o}{f}\right]  \text{mK}^2
\end{split}
\end{equation}
where $k$ is the magnitude of the $k$-mode, $B$ is the bandwidth, $\Delta\ln k$ is the log
of the binsize, $\Omega$ is the field-of-view, $T_{\text{sys}}$ is the system temperature, 
${t_\text{per\_day}}$ is the number of hours observed per day, $t$ is the number of days
observed, $N$ is the number of antennas, and $f_o/f$ is the configuration metric for a 
redundant array as defined in Parsons et al.

Note that $\Delta^2_N(k)$ is the standard radiometric sensitivity equation, scaled by
the volume in $k$-space, normalized by the power spectrum Fourier coefficient, and
reduced by the number of independent samples in a given $k$-mode bin, which may have
both coherent and incoherent application.

Pulling out terms relating to diameter and number, we can write Eq. \ref{eq:sensitivity} as
\begin{equation}
\label{eq:reducedSensitivity}
\Delta^2_N(k) \propto \frac{\Omega (f_o/f)}{N\sqrt{t_\text{per\_day}}} \propto \frac{(1/D^2)(1/\sqrt{N})}{N\sqrt{D}}
= D^{-\frac{5}{2}}N^{-\frac{3}{2}}
\end{equation}
where the dependencies on diameter and number have been subsituted in, noting that the expressions for $f_o/f$ and 
$t_\text{per\_day}$were derived in Parsons et al where the baselines for the close-packed array are multiples of the diameter.

For a fixed sensitivity, we may then parameterize the performance/cost ratio by the diameter.  Figure \ref{fig:costing}

\section{System Overview}
\label{sec:design}

We have achieved a
pivotal new understanding of how instrumental characteristics give rise to the
wedge of emission shown in Figure \ref{fig:twoFGViews}.  Furthermore, we have
measurements that prove the efficacy, to the sensitivity limits of current
instruments, of projecting out foreground-dominated wavemodes within the
wedge.  The HI cosmology community is now in a position to define
the top-level instrument requirements that ensure foregrounds remain bounded
within the wedge, and to specify the sensitivity needed to obtain
high-significance detections of the 21~cm reionization power spectrum under the
conservative assumption that all foreground-dominated wavemodes within the
wedge must be projected out of our measurements.
As summarized in Table \ref{tab:signif} (see \Mycitealt{pober_et_al2014} for more details),
based on these assumptions, current instruments are likely to achieve,
at best, only marginal detections of power from reionization.

%HERA follows the vision delineated in NWNH where the MWA and PAPER
%characterized the foreground emission and performed the first deep
%integrations, followed by a more powerful instrument that drew on the lessons
%learned and merged the scientific teams. The MWA, PAPER, and LOFAR have all
%recently started deep integrations with hopes of detecting the 21~cm
%signal.
%However, as summarized in Table \ref{tab:signif} (see \citealt{pober_et_al2014} for more details),
%these instruments are likely to achieve
%at best only marginal detections of power from reionization.
%While this might qualify as a ``detection'' if the signal is strong, the
%\emph{science} of reionization observations is beyond their reach. 
%However, the profound contributions of MWA and PAPER should not be overlooked. PAPER and the MWA have succeeded magnificently in characterizing the foreground emission, discovering the EoR window, and understanding how to remove foreground contamination from the measurements. Further, they have developed and retired the risk associated with all of HERA's major systems. 
%HERA uses these lessons perform the science of reionization, at a much lower cost than assumed in NWNH.
%% that this proposal is cheaper than NWNH suggested but never give either that
%% number nor your estimated cost.

This proposal targets two arrays: a 127-element array that borrows heavily
from components of the PAPER experiment, and an upgraded 331-element array that
incorporates several performance optimizations.  
These arrays are developed over four years (see Fig. \ref{fig:timeline} and the Project Management Plan for details),
with recurring cycles of development, testing, review, deployment, commissioning, and observation.
As listed in Table \ref{tab:signif}, the specifications of HERA-127
and HERA-331 have been set to meet precisely the requirements for obtaining,
respectively, a 10$\sigma$ detection of the 21~cm reionization signal across a broad range of redshifts, and a
25$\sigma$ detection of the power spectrum of fluctuations in the 21~cm signal
capable of determining the nature and distribution of the first galaxies that
dominate cosmic reionization.  
As discussed in
\S\ref{sec:Lessons}, these science requirements translate directly into
requirements for signal delay and collecting area that, when 
combined with a cost minimization requirement, tightly
bound the HERA design.  The basic parameters are given in Table \ref{tab:BasicParameters}.

%incorporates
%our proven foreground avoidance techniques while improving
%dramatically the sensitivity relative to current experiments.  With a
%new understanding of how antenna size and separation affect
%sensitivity and foreground isolation, it has become evident a revision
%of the PAPER antenna design can yield up to 20 times the sensitivity
%per element without substantially degrading foreground isolation.
%Where PAPER's elements lack collecting area and are smaller than
%strictly required for foreground isolation, and the majority of MWA
%and LOFAR elements are spaced too widely to avoid foregrounds,
%HERA-331 employs an extremely compact array of 14-m parabolic dishes
%with PAPER-style dipole feeds (see Figure \ref{fig:hera_dish}).  The dishes
%are placed in a compact hexagonal configuration of 331 dishes supplemented by 18 outrigger dishes to achieve dense and
%highly-redundant $uv$-sampling (see Figures \ref{fig:uv_coverage} and \ref{fig:config_optics}).  The dishes have a short (4.5m) focal height to limit the
%path length of reflections, whose time-delay gives rise to chromatic
%instrumental systematics.
%
%
%% AAARRGGHH!
%%\begin{figure}[t]
%%\centering
%%	\begin{subfigure}[b]{0.46\textwidth}
%%		\includegraphics[width=\textwidth]{plots/paper_element.jpg}
%%		\Caption{-0.2in}{0.99}{-0.2in}{The PAPER element (provides a clean instrumental response as a function
%%		of frequency \citep{parsons_et_al2010,parsons_et_al2012b}, which is crucial to
%%		the foreground isolation shown in Figure \ref{fig:eor_pspec}.}
%%	\end{subfigure}
%%	\quad
%%	\begin{subfigure}[b]{0.46\textwidth}
%%		\includegraphics[height=1.75in]{plots/hera_dish.png}
%%		\Caption{-0.2in}{0.99}{-0.2in}{A 14m dish designed around the feed dramatically improves sensitivity while
%%		constraining the path length and amplitude of reflections to ensure that foreground 
%%		isolation is not substantially degraded.}
%%	\end{subfigure}
%%\Caption{-0.2in}{0.99}{-0.2in}{PAPER and HERA elements}
%%\label{fig:hera_dish}
%%\end{figure}
%
%The size of HERA-331 dishes optimizes cost for a fixed sensitivity and
%level of foreground isolation.  The associated reduction in the number
%of antenna elements to achieve a given collecting area, combined with
%the fact that these dishes have no moving parts, are built from
%inexpensive materials, and follow a simple construction that can be
%contracted locally, makes the cost of building HERA-331 substantially
%cheaper than was anticipated in the roadmap submitted to \nwnh\ for this
%stage of the program.   
%
%HERA leverages the technical heritage of PAPER, MWA and of CASPER\footnote{Collaboration for Astronomical
%Signal Processing and Electronics Research, a Berkeley-initiated worldwide open source community that is
%developing boards, firmware and software for the astronomical community} and incorporates a
%phased implementation to mitigate against risk.  The system, technology and phased approach is discussed below
%in more detail.
%The HERA instrument design is very straightforward.  The element
%itself is a fixed zenith-pointing 14-meter segmented prime-focus paraboloid with a high screen
%to minimize cross-talk between elements (which are spaced 14.3 meters on a
%hexagonal grid).   The $f/D$ of
%the paraboloid is 0.32, so that the focal length, $f$, is less than 5 meters to meet the 
%standing wave specification at the delays of interest of more than 60 dB of attenuation at delays 
%greater than 50 ns.
%
%
%The active feed sends back the entire dual-polarization analog bandwidth on standard
%coaxial cable to an aggregation point called a ``node'', which services 
%about 15 antennas.  This cable length is kept short (35-m) to keep any standing
%wave contamination outside of the delay-space of interest for power spectrum
%measurements.  The node amplifies, filters, digitizes and transmits the signal data stream
%back to the central processing location (the Karoo Array Processing Building - KAPB).
%Figure \ref{fig:blockDiagram} shows a block diagram of the system.

As shown in Figure \ref{fig:systemOverview}, the HERA instrument has a straightforward signal
path from antenna elements
with active feeds, to nodes that digitize and aggregate signals onto an optical network,
and on to a processing building where signals are correlated and data are stored, compressed,
and shipped to a computing cluster located at U. Pennsylvania (UPenn).

\subsection{Antenna Element}

The novel design of HERA's antenna element
(Fig. \ref{fig:hera_dish}, right panel) represents a critical advance
that enables HERA to achieve its science goals.  This 14-m
fixed zenith-pointing parabolic dish yields more than 10 times the
sensitivity of an MWA tile (and more than 20 times
that of a PAPER element), but does so without substantially degrading
our ability to isolate and remove foreground emission on the basis of
spectral smoothness.  As described above, this is done by limiting
the timescale of signal delays and reflections to under 120 ns.

Since the time it takes radio emission to propagate between adjacent antennas 
represents a significant portion of this time budget, HERA elements must be placed close together.
Recent work characterizing foregrounds suggests that 
antenna separations of $8\lambda \approx 15$m are
well-behaved to current limits \Mycitep{pober_et_al2013,parsons_et_al2013}. This influences our
choice of a 14-m dish diameter,
which incurs 42 ns of signal delay between adjacent antennas.
The focal height (4.5 m) derives from the illumination pattern of the prime-focus
feed and the fact that reflections
between the feed and the element
%caused by imperfect impedance matching between the feed electronics and free space, 
can introduce additional signal delay. 
%With a splash cone underneath the feed to scatter reflections, we assume -20 dB
%attenuation per reflection, so that after three traversals (13.5 m; 41 ns), reflections from
%foreground emission are below the expected level of the 21~cm reionization signal.
Additional measures, such as a splash cone underneath the feed, screens that isolate feeds from one another, and 
non-metallic cabling and poles for supporting the feeds, are all aimed at minimizing
the potential for additional sources of signal reflections.
The total signal delay associated with the HERA dish design is estimated at 83 ns, leaving
headroom for reflections arising in the analog signal path.

Cost is another design constraint, including the price of construction
materials, assembly in a remote area, and maintenance over the operational
lifetime of the array.  Care has been taken to select robust and inexpensive
construction materials (PVC, concrete, utility poles, 0.25" wire cloth) and an assembly methodology that delivers the required positional
accuracy of 10 cm and surface accuracy of 2 cm given standard expertise in construction
practices for the subcontracted teams that set the poles and construct the
elements in the field.  
%JCP: "required positional accuracy" -- what drives the requirement?
Given the sensitivity and signal delay requirements, the size of HERA
dishes optimizes a global costing curve that includes the costs of the elements,
the signal path, correlation, data storage, and processing.

% ARP: removing this figure, since it appears in project management plan
%\begin{figure}[h]
%	\centering
%		\includegraphics[width=0.46\textwidth]{plots/heracles.png}
%		\includegraphics[width=0.46\textwidth]{plots/Engineering/heraclesNA.png}
%	\caption{\small
%		The construction prototype element in California (left) with preliminary 
%        reflection measurements (right).
%} \label{fig:heracles}
%\end{figure}

%As illustrated in Figure \ref{fig:heracles}, 
A first prototype of the HERA dish is nearly completely constructed.  In this
proposal, UC Berkeley and NRAO 
lead the incorporation of lessons from the
construction process and reflectometry tests into the design.
Two further prototype elements are constructed in the first year alongside the existing
PAPER array in Green Bank, WV,
for on-sky measurements of the beam pattern and spectral 
response.
These measurements are coupled with full electromagnetic modeling to test
and refine the element design prior to a Critical Design Review at the end of the first project
year.  Three revised elements are then constructed in South Africa, involving
lead members of the construction subcontract teams for a Project Readiness Review.  %Thereafter,
%element construction will proceed continously through the second and third project years during the day,
%with a staged incorporation of new elements for night-time observing.

%The elements are
%also nearly abutting one another, which allows for sharing of physical support
%infrastructure.

% ARP: lack room for this figure.
%\begin{figure}[h]
%	\centering
%	\begin{subfigure}[b]{0.46\textwidth}
%		\includegraphics[width=0.99\textwidth]{plots/Engineering/nvsd.png}
%		\caption{Costing model for a fixed sensitivity and varying the diameter.   The arrows indicate the regions of 
%				increasing systematics and delay-space contamination.}
%		\label{fig:nvsd} 
%	\end{subfigure}
%	\quad
%	\begin{subfigure}[b]{0.46\textwidth}
%		\includegraphics[width=0.8\textwidth]{plots/Engineering/focalEff.png}
%		\caption{Analytical model efficiency of a parabolic element as a function of focal height and diameter.   The delay 
%				contamination specification is $f<5$m.}
%		\label{fig:disheffic}
%	\end{subfigure}
%	\caption{Cost and performance variational analysis data for the element.}
%\end{figure}

% FOLLOWING SEEMS TOO DETAILED FOR A PROPOSAL
%To distinguish between measured sky delays and instrumental-induced delays at a required threshold ($R_T$) dB, 
%we need sufficient attenuation of the reflected signals at a given focal length ($f$) at a specified delay ($\tau_{d}$).  Assuming a conservative simple model that the magnitude of the reflection is attenuated by $A$ dB at each reflection 
%we see that for a delay length limit of $\delta_{d}=c\tau_{d}$ and focal length $f$,  the required focal length is
%
%\begin{equation}
%f < \left(\frac{A}{R_T}\right)\delta_d.
%\end{equation}
%
%Using nominal values of $R_T$ =  60dB (an order of
%magnitude below where EoR is predicted to be below foregrounds) at delays
%corresponding to the time it takes travel 15m and a net attenuation of 20 dB per reflection, 
%we find that the focal length should be less than about 5 m.  Free-space loss effects would 
%increase that value, loosening the constraint.
%DETAILEND

%In addition to the constraints given by the element itself, the HERA element size is
%also influenced by the location of the ``knee'' in the EoR power spectrum \citep{lidz_et_al2008}
%The EoR power spectrum has an upward slope for low $k$-modes which levels off
%around $k=0.15 h$/Mpc %(see fig blah). 
%Working inside this $k$-mode would be beneficial due to the fact foregrounds are
%less problematic. This poses a problem because without knowing the
%width of our foregrounds, we can't say for sure which $k$-modes (in the power
%spectrum) are corrupted. This uncertainty, coupled with increasing systematic affects for shorter 
%baselines (hence smaller diameters), favors larger diameter antennas.

%% ARP: no room for this figure.
%\begin{figure}[h]
%	\centering
%	\begin{subfigure}[b]{0.46\textwidth}
%		\includegraphics[width=0.85\textwidth]{plots/dish.png}
%		\caption{CAD model of 14m dish with screening and some supports removed to show detail.}
%		\label{fig:dish} 
%	\end{subfigure}
%\quad
%	\begin{subfigure}[b]{0.46\textwidth}
%		\includegraphics[width=\textwidth]{plots/Engineering/hera_beam.png}
%		\caption{Analytical model beam patterns at 10m, 12m and 14m.}
%		\label{fig:beam} 
%	\end{subfigure}
%	\caption{HERA element and beam pattern.}
%\end{figure}

%The $k$-mode for a $14m$ baseline is $k = 0.023$, given by $k_{H} =
%\frac{B}{c}\frac{dk}{d\eta}$, where $B$ is the length of the baseline, 14m in
%our case, $c$ is the speed of light, and $\frac{dk}{d\eta}$ is the cosmological
%transfer function from delays to $k$-modes. In addition, narcissistic
%reflections add into our $k$ budget as well. The $f$ and $D$ choices above provide a $k$ budget for
%foreground widths to be within $\Delta{k}\sim{0.1}$. Testing these hypothesis
%and again, finding a compromise between maximum sensitivity, foreground budget
%will be key to testing and constructing the required element. Foreground width constraints
%are still an area of active research \citep{pober_et_al2013}.

% The core defining elements in this design are the central hub and three tall support poles.  Three intermediate 
% support posts are installed between each pair of poles.  A 2$^{\prime\prime}$ PVC spar of 24.1$^{\prime}$ 
% terminates at each pole and post.  These spars are supported at each end and one point in the middle at the 
% proper height and angle.  The intervening PVC pipe essentially acts as a smoothing filter between those 
% points, noting also that a beam with point loads attains nearly the quadratic shape desired.  The CAD model is 
% shown in the right panel of Figure \ref{fig:hera_dish}.  
% 
% The tall ($\sim$ 7m) poles provide locational accuracy (in all three dimensions) for the overall array installation.  
% Using conventional commercial pole-installation techniques the poles are installed first for the entire array.  
% Note that every pole except for the edge poles are shared by three antennas.  A standard theodolite can 
% then be used to mark a known level height on all three poles.  These locations are then used with tensioned 
% lines to define the center of that element and the hub is positioned at that location using a jig.
% 
% The hub uses concentric commercially available ``sonotube'' forms (circular cardboard forms for concrete pillars) 
% and PVC sleeves to hold the PVC spars and PVC supports.  The retaining holes may be accurately cut into 
% the forms, sleeves installed and concrete poured to make a simple hub to the desired accuracy.  The jig holds 
% the concentric rings in place and allows it to be centered by tensioned lines while the concrete is poured.  
% When the concrete cures one can then transfer an accurate offset from the dish vertex back to the poles.
% The intermediate posts are then located by the support sub-assemblies attached to the poles and posts along 
% with the rim sub-assemblies.  These are positioned and a small pier is poured to locate them.  After spar support 
% pieces are installed on the posts, the spars themselves and the metal cloth can be installed.
% 
% The feed is held off the three tall poles using tensioned lines to accurately locate it over the hub.  A precise 
% length of kevlar rope holds the feed down to the hub at a precise focal point.  The RF cables follow the line 
% down and out to the analog-to-digital converters and correlator.  
% 
% To minimize cross-talk, metal screens are strung between every pole/post, which go to the level of the feed.  
% The dishes have a rim-to-rim spacing of 30cm to allow the screen to be slightly angled to minimize standing waves.

\subsection{Array Configuration}

 Figure \ref{fig:config}

For HERA-127 and HERA-331, elements are arranged in a compact hexagonal grid.
This configuration minimizes antenna separation, which is critical for meeting
the science requirements for foreground isolation, and 
produces a highly redundant sampling of the $uv$ plane.  As described in
\S\ref{sec:Lessons}, redundant configurations have been employed by PAPER to
boost sensitivity \Mycitep{parsons_et_al2012a}, with the added benefit that they
greatly facilitate fast and accurate calibration
\Mycitep{liu_et_al2010,parsons_et_al2013}.  Placing HERA elements in a
grid allows certain construction components to be re-used between antennas,
reducing cost and improving the accuracy of element placement.  HERA-331 
is a build-out of HERA-127 from the eastern edge.

The HERA-331 core has excellent
imaging capability that can be leveraged for developing foreground suppression techniques
to improve access to the 21~cm reionization signal 
(Fig. \ref{fig:eor_pspec}, black).
This capability is augmented with 21 outrigger elements.  These outriggers
combine with the dense core to generate a fully
sampled $uv$ plane out to 400$\lambda$.  While outriggers on the southern half of the array are
placed in alignment with the hexagonal grid in the core, elements on the northern side are
shifted off-grid to sample the $uv$ plane at sub-aperture scales.
This sampling strategy helps eliminate grating lobes in the
synthesized beam and provides information for
calibrating and correcting direction-dependent antenna responses.  This capability may
be used to match polarization beams and minimize
polarization leakage.

\subsection{Analog Signal Path}
